---
title: ""
author: "Tino Schneidewind"
date: "2025-03-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r libs, echo = FALSE, warning=FALSE, message=FALSE}
library(dplyr); library(tidyverse); library(ggplot2); library(grid); library(gridExtra)
library(readxl);library(knitr);library(lubridate); library(readr);library(zoo);library(fuzzyjoin)

rm(list = ls())
# source("../MACorWIN.R")
```

## Data Cleaning Protocol (revised)

The aim of this document is to derive and present a method of multiple steps to be taken that convert the raw *personal environmental data* into an analysis ready dataset. For this I selected the data of 5 participants: *ACT001D* (very good data from visual inspection), *ACT004S* and *ACT014F* (some poor data), *ACT003C* and *ACT032V* (very poor data). The variables to be cleaned are temperature, relative humidity RH, and noise.

Main reasons for cleaning are:

- devices were taken off person
- the device malfunctioned
- issues in data processing from device to data

The steps of cleaning after revision are the following: 

1. Study design limitations (extracted from PVL)                           
2. Physically possible (temperature > 0K, noise > 0 dB, RH < 100% etc).            
3. Physically plausible (temperature < -10 °C etc) **especially important for taped temperature filtering**
4. Variability (variable specific threshold of standard deviation over an extended period of time) **especially important for worn temperature/humidity**

```{r load redcap, echo = FALSE, warning=FALSE, message=FALSE}
# REDCap for uids and start and end times
# redcap = read_csv("/Volumes/FS/_ISPM/CCH/Actual_Project/data/App_Personal_Data_Screening/redcap_data.csv")
redcap <- read_csv("../../redcap_all.csv")

redcap_pvl <-  redcap |>
  select( uid, redcap_event_name, pvl_start, pvl_end,pvl_ibuthouse, pvl_ibutworn, pvl_ibuttaped) |>
  filter(!is.na(pvl_start)) |>
  filter(uid %in% c("ACT057R", "ACT028W", "ACT016N", "ACT067C", "ACT043W"))  |>
  mutate(redcap_event_name = substr(redcap_event_name, 13,18)) |>
  filter(redcap_event_name == "week_2")


```


```{r data, message=FALSE, warning=FALSE, echo=FALSE}
# uncleaned but selected by pvl
data = read_csv("/Volumes/FS/_ISPM/CCH/Actual_Project/data/App_Personal_Data_Screening/week_2_minute_data_unclean.csv") |>
   filter(uid %in% c("ACT057R", "ACT028W", "ACT016N", "ACT067C", "ACT043W"))  |>
    mutate(Variable = str_replace(Variable, "_NS", "NS"))

# house data
data_H <- data |>
  filter(Variable == "IBH_HUM" | Variable == "IBH_TEMP")|>
  pivot_wider(names_from = Variable, values_from = Value)|>
  mutate(
    IBH_HUM_MA = rollmean(IBH_HUM, k = 8, fill = NA, align = "left"),
    IBH_TEMP_MA = rollmean(IBH_TEMP, k = 8, fill = NA, align = "left"),
  )

# worn data
data_W <- data |>
  filter(Variable == "IBW_HUM" | Variable == "IBW_TEMP")|>
  pivot_wider(names_from = Variable, values_from = Value)|>
  mutate(
    IBW_HUM_MA = rollmean(IBW_HUM, k = 8, fill = NA, align = "left"),
    IBW_TEMP_MA = rollmean(IBW_TEMP, k = 8, fill = NA, align = "left"),
    IBW_HUM_MSD = rollapply(IBW_HUM, width = 3, FUN = sd, align = "left", fill = NA)
  )

# taped data
data_T <- data |>
  filter(Variable == "IBT_TEMP")|>
  pivot_wider(names_from = Variable, values_from = Value)|>
  mutate(
    IBT_TEMP_MA = rollmean(IBT_TEMP, k = 8, fill = NA, align = "left"),
    IBT_TEMP_MSD = rollapply(IBT_TEMP, width = 3, FUN = sd, align = "left", fill = NA)
  )

# noise data
data_N <- data |>
  filter(Variable == "NS")|>
  pivot_wider(names_from = Variable, values_from = Value)|>
  mutate(
    NS_MA = rollmean(NS, k = 8, fill = NA, align = "left")
  )

# summary table for variability
summary_manual <- data.frame(
  Variable = c("IBH_HUM", "IBH_TEMP", "IBW_HUM", "IBW_TEMP", "IBT_TEMP", "NS"),  
  mean = c(mean(data_H$IBH_HUM, na.rm = TRUE), mean(data_H$IBH_TEMP, na.rm = TRUE),
           mean(data_W$IBW_HUM, na.rm = TRUE), mean(data_W$IBW_TEMP, na.rm = TRUE),
           mean(data_T$IBT_TEMP, na.rm = TRUE), mean(data_N$NS, na.rm = TRUE)),
  sd = c(sd(data_H$IBH_HUM, na.rm = TRUE), sd(data_H$IBH_TEMP, na.rm = TRUE),
         sd(data_W$IBW_HUM, na.rm = TRUE), sd(data_W$IBW_TEMP, na.rm = TRUE),
         sd(data_T$IBT_TEMP, na.rm = TRUE), sd(data_N$NS, na.rm = TRUE)))
```




### 1. Study Design: Observation period and excluding PVL-visits

First, data has to be excluded that was taken outside the observation window  and during personal visit log times **if the devices were changed**. The data was cut the the observation window in th data compiling but the checking whether the device was changed will be done here. No additional values were excluded from the chosen 5 individuals.



```{r filter for device changes, echo = FALSE, warning=FALSE, message=FALSE}
# loop through uids
for(uids in unique(redcap_pvl$uid)){

  redcap_subset <- redcap_pvl[redcap_pvl$uid == uids,]
  
  # loop through pvl visits
  for(i in nrow(redcap_subset)) {
    
    
      startvalue <- redcap_subset$pvl_start[i]
      endvalue <-  redcap_subset$pvl_end[i]
    
    # if house Ibutton was changed
    if(redcap_subset$pvl_ibuthouse[i] > 0 ){
      
      # set values to NA
      data_H <- data_H |>
        mutate(IBH_HUM = if_else(uid %in% uids & datetime > startvalue & datetime < endvalue, NA, IBH_HUM),
               IBH_TEMP = if_else(uid %in% uids & datetime > startvalue & datetime < endvalue, NA, IBH_TEMP),
               IBH_HUM_MA = if_else(uid %in% uids & datetime > startvalue & datetime < endvalue, NA, IBH_HUM_MA),
               IBH_TEMP_MA = if_else(uid %in% uids & datetime > startvalue & datetime < endvalue, NA, IBH_TEMP_MA))
    }
      
    # if worn Ibutton was changed
    if(redcap_subset$pvl_ibutworn[i] > 0 ){
      
      # set values to NA
      data_W <- data_W |>
        mutate(IBW_HUM = if_else(uid %in% uids & datetime > startvalue & datetime < endvalue, NA, IBW_HUM),
               IBW_TEMP = if_else(uid %in% uids & datetime > startvalue & datetime < endvalue, NA, IBW_TEMP),
               IBW_HUM_MA = if_else(uid %in% uids & datetime > startvalue & datetime < endvalue, NA, IBW_HUM_MA),
               IBW_TEMP_MA = if_else(uid %in% uids & datetime > startvalue & datetime < endvalue, NA, IBW_TEMP_MA),
               IBW_HUM_MSD = if_else(uid %in% uids & datetime > startvalue & datetime < endvalue, NA, IBW_HUM_MSD))
    }
    
    # if taped Ibutton was changed
    if(redcap_subset$pvl_ibuttaped[i] > 0 ){
      
      # set values to NA
      data_T <- data_T |>
        mutate(IBT_TEMP = if_else(uid %in% uids & datetime > startvalue & datetime < endvalue, NA, IBT_TEMP),
               IBT_TEMP_MA = if_else(uid %in% uids & datetime > startvalue & datetime < endvalue, NA, IBT_TEMP_MA),
               IBT_TEMP_MSD = if_else(uid %in% uids & datetime > startvalue & datetime < endvalue, NA, IBT_TEMP_MSD))
    }
  }
}
```

```{r study design plots, fig.align='center', fig.width=12, fig.height=2.5, echo=FALSE}
for (uids in unique(data$uid)) {
  
  plot_H <- data_H |>
    filter(uid == uids) |>
  ggplot(aes(x = datetime)) +
    geom_line(aes(y = IBH_HUM), color = "skyblue2") +
    geom_line(aes(y = IBH_TEMP), color = "brown2") +
    theme_classic() +
    lims(y=c(15,80)) +
    labs(x = "time", y = "value", title = paste0(uids, " - HOUSE"))
  
  plot_W <- data_W |>
    filter(uid == uids) |>
  ggplot(aes(x = datetime)) +
    geom_line(aes(y = IBW_HUM), color = "skyblue2") +
    geom_line(aes(y = IBW_TEMP), color = "brown2") +
    theme_classic() +
    lims(y=c(15,80)) +
    labs(x = "time", y = "value", title = paste0(uids, " - WORN"))
  
  plot_T <- data_T |>
    filter(uid == uids) |>
  ggplot(aes(x = datetime, y = IBT_TEMP)) +
    geom_line(color = "brown2") +
    theme_classic() +
    labs(x = "time", y = "value", title = paste0(uids, " - TAPED"))
  
    plot_N <- data_N |>
    filter(uid == uids) |>
  ggplot(aes(x = datetime, y = NS)) +
    geom_line(color = "grey2") +
    theme_classic() +
    labs(x = "time", y = "value", title = paste0(uids, " - Noise"))
  
    grid.arrange(plot_H, plot_W, plot_T, plot_N, ncol=4)
}
```

<br> 

### 2. Physically possible

Every Variable (temperature, RH, noise) has its physical limits that the following:

1. Temperature: < -273 °C
2. RH: < 0 % and > 100 %
3. Noise: < 0 dB

```{r physically possible}
# House
data_H <- data_H |>
  mutate(IBH_TEMP_01 = if_else(IBH_TEMP < -273, 1, 0),
         IBH_HUM_01 = if_else(IBH_HUM < 0 | IBH_HUM > 100, 1, 0))

# Worn
data_W <- data_W |>
  mutate(IBW_TEMP_01 = if_else(IBW_TEMP < -273, 1, 0),
         IBW_HUM_01 = if_else(IBW_HUM < 0 | IBW_HUM > 100, 1, 0))

# Taped
data_T <- data_T |>
  mutate(IBT_TEMP_01 = if_else(IBT_TEMP < -273, 1, 0))

# Noise
data_N <- data_N |>
  mutate(NS_01 = if_else(NS < 0, 1, 0))

```

No plots are shown here because there are no impossible values in the example data.


```{r plot physically possible, fig.align='center', fig.width=12, fig.height=2.5, echo=FALSE, message=FALSE, warning=FALSE }
# for (uids in unique(data$uid)) {
#   
#   plot_H <- data_H |>
#     filter(uid == uids) |>
#     ggplot(aes(x = datetime)) +
#     geom_line(aes(y = IBH_HUM), color = "skyblue2") +
#     geom_line(aes(y = IBH_TEMP), color = "brown2") +
#     geom_point(aes(y = ifelse(IBH_HUM_01 == 1 & !is.na(IBH_HUM_01), IBH_HUM, NA_real_)), color = "blue") +
#     geom_point(aes(y = ifelse(IBH_TEMP_01 == 1 & !is.na(IBH_TEMP_01), IBH_TEMP, NA_real_)), color = "brown") +
#     theme_classic() +
#     lims(y = c(15, 80)) +
#     labs(x = "time", y = "value", title = paste0(uids, " - HOUSE"))
#   
#   plot_W <- data_W |>
#   filter(uid == uids) |>
#   ggplot(aes(x = datetime)) +
#     geom_line(aes(y = IBW_HUM), color = "skyblue2") +
#     geom_line(aes(y = IBW_TEMP), color = "brown2") +
#     geom_point(aes(y = ifelse(IBW_HUM_01 == 1 & !is.na(IBW_HUM_01), IBW_HUM, NA_real_)), color = "blue") +
#     geom_point(aes(y = ifelse(IBW_TEMP_01 == 1 & !is.na(IBW_TEMP_01), IBW_TEMP, NA_real_)), color = "brown") +
#     theme_classic() +
#     lims(y = c(15, 80)) +
#     labs(x = "time", y = "value", title = paste0(uids, " - WORN"))
#   
#   plot_T <- data_T |>
#     filter(uid == uids) |>
#   ggplot(aes(x = datetime, y = IBT_TEMP)) +
#     geom_line(color = "brown2") +
#     geom_point(aes(y = ifelse(IBT_TEMP_01 == 1 & !is.na(IBT_TEMP_01), IBT_TEMP, NA_real_)), color = "brown") +
#     theme_classic() +
#     labs(x = "time", y = "value", title = paste0(uids, " - TAPED"))
#   
#     plot_N <- data_N |>
#     filter(uid == uids) |>
#     mutate(NS_05 = replace_na(NS_05, 0)) |>  # Ensure no NA in NS_01
#     ggplot(aes(x = datetime)) +
#     geom_line(aes(y = NS), color = "grey") +
#     geom_point(aes(y = ifelse(NS_01 == 1 & !is.na(NS_01), NS, NA_real_)), color = "black") +  # Use NA_real_
#     theme_classic() +
#     labs(x = "time", y = "value", title = paste0(uids, " - Noise"))
#   
#     grid.arrange(plot_H, plot_W, plot_T, plot_N, ncol=4)
# }
```

<br>

### 3. Physically plausible

The plausible range is to some degree subjective, depends on the observation surroundings and changes not only depending on the variable, but also what the variable describes (temperature taped and house). Therefore now we need to start with device specific variable value ranges. 

1. House: Temperature < 0 °C and > 55 °C, RH: no additional filtering
2. Worn: Temperature < 10 °C and > 45 °C, RH: no additional filtering
3. Taped: Temperature below the 10th percentile (no upper filtering because taped temperature almost always is greater than house temperature) (25th percentile was too high)
4. Noise: no additional filtering


```{r physically plausible}
# House
data_H <- data_H |>
  mutate(IBH_TEMP_02 = if_else(IBH_TEMP < 0 | IBH_TEMP > 55, 1, 0))

# Worn
data_W <- data_W |>
  mutate(IBW_TEMP_02 = if_else(IBW_TEMP < 15 | IBW_TEMP > 45, 1, 0))

# Taped

# IQR 
Q25 = quantile(data_T$IBT_TEMP, .10)
data_T <- data_T |>
  mutate(IBT_TEMP_02 = if_else(IBT_TEMP < Q25, 1, 0))
```


```{r plot physically plausible, fig.align='center', fig.width=12, fig.height=2.5, echo=FALSE, message=FALSE, warning=FALSE }
for (uids in unique(data$uid)) {
  
  plot_H <- data_H |>
    filter(uid == uids) |>
    ggplot(aes(x = datetime)) +
    geom_line(aes(y = IBH_HUM), color = "skyblue2") +
    geom_line(aes(y = IBH_TEMP), color = "brown2") +
    # geom_point(aes(y = ifelse(IBH_HUM_05 == 1 & !is.na(IBH_HUM_05), IBH_HUM, NA_real_)), color = "blue") +
    geom_point(aes(y = ifelse(IBH_TEMP_02 == 1 & !is.na(IBH_TEMP_02), IBH_TEMP, NA_real_)), color = "brown") +
    theme_classic() +
    lims(y = c(15, 80)) +
    labs(x = "time", y = "value", title = paste0(uids, " - HOUSE"))
  
  plot_W <- data_W |>
  filter(uid == uids) |>
  ggplot(aes(x = datetime)) +
    geom_line(aes(y = IBW_HUM), color = "skyblue2") +
    geom_line(aes(y = IBW_TEMP), color = "brown2") +
    # geom_point(aes(y = ifelse(IBW_HUM_05 == 1 & !is.na(IBW_HUM_05), IBW_HUM, NA_real_)), color = "blue") +
    geom_point(aes(y = ifelse(IBW_TEMP_02 == 1 & !is.na(IBW_TEMP_02), IBW_TEMP, NA_real_)), color = "brown") +
    theme_classic() +
    lims(y = c(15, 80)) +
    labs(x = "time", y = "value", title = paste0(uids, " - WORN"))
  
  plot_T <- data_T |>
    filter(uid == uids) |>
  ggplot(aes(x = datetime, y = IBT_TEMP)) +
    geom_line(color = "brown2") +
    geom_point(aes(y = ifelse(IBT_TEMP_02 == 1 & !is.na(IBT_TEMP_02), IBT_TEMP, NA_real_)), color = "brown") +
    theme_classic() +
    labs(x = "time", y = "value", title = paste0(uids, " - TAPED"))
  
    plot_N <- data_N |>
    filter(uid == uids) |>
    # mutate(NS_05 = replace_na(NS_05, 0)) |>  # Ensure no NA in NS_01
    ggplot(aes(x = datetime)) +
    geom_line(aes(y = NS), color = "grey") +
    # geom_point(aes(y = ifelse(NS_05 == 1 & !is.na(NS_05), NS, NA_real_)), color = "black") +  # Use NA_real_
    theme_classic() +
    labs(x = "time", y = "value", title = paste0(uids, " - Noise"))
  
    grid.arrange(plot_H, plot_W, plot_T, plot_N, ncol=4)
}
```

<br>

### 4. Variability 

The variability between variables and devices differs significantly (eg. humidity house and worn). Because we are interested in stress experienced by the individuals, it is important to not filter out extreme but realistic conditions as these represent the largest stress impact. However we do want to filter out worn measurements that resemble the variance of the house measurements and indicate the the device was not worn. We use the moving standard deviation of 3 left aligned humidity values. As an additional measure to prevent filtering out reasonable values, we filter only measurements if the standard deviation has been too low for 2 consecutive measurements. We initially considered 4 values, however, in the averaging process this low threshold then introduced "wrong" data back into the hourly averages.

1. Worn: threshold for humidity sd: x

```{r variance}
x = 1

# Worn
data_W <- data_W |>
  mutate(IBW_TEMP_04_intermediate = if_else(IBW_HUM_MSD < x, 1, 0),
         IBW_HUM_04_intermediate = if_else(IBW_HUM_MSD < x, 1, 0),
         IBW_TEMP_04 = rollmean(IBW_TEMP_04_intermediate, k = 2, fill = NA, align = "left"),
         IBW_HUM_04 = rollmean(IBW_HUM_04_intermediate, k = 2, fill = NA, align = "left"))
```

```{r variance plot, fig.align='center', fig.width=12, fig.height=5, echo=FALSE, message=FALSE, warning=FALSE }
plot_list <- list()

for (i in 1:length(unique(data$uid))) {
  
  uids <-  unique(data$uid)[i]
  
  plot_W <- data_W |>
    filter(uid == uids) |>
    ggplot(aes(x = datetime)) +
    geom_line(aes(y = IBW_HUM), color = "skyblue2") +
    geom_line(aes(y = IBW_TEMP), color = "brown2") +
    geom_point(aes(y = ifelse(IBW_HUM_04 == 1 & !is.na(IBW_HUM_04), IBW_HUM, NA_real_)), color = "blue") +
    geom_point(aes(y = ifelse(IBW_TEMP_04 == 1 & !is.na(IBW_TEMP_04), IBW_TEMP, NA_real_)), color = "brown") +
    theme_classic() +
    lims(y = c(15, 80)) +
    labs(x = "time", y = "value", title = paste0(uids, " - WORN"))

    plot_list[[i]] <- plot_W
}

grid.arrange(grobs = plot_list, ncol = 4)
```

<br>

### Combination of all cleaning methods

The plots below show the cleaned data including all cleaning methods. Light colors indicate cleaned original data.

```{r final, echo=FALSE, warning=FALSE, message=FALSE}
# House
data_H <- data_H |>
  mutate(IBH_HUM_all = rowSums(across(c(IBH_HUM_01)), na.rm = TRUE),
         IBH_TEMP_all = rowSums(across(c(IBH_TEMP_01, IBH_TEMP_02)), na.rm = TRUE))

# Worn
data_W <- data_W |>
  mutate(IBW_HUM_all = rowSums(across(c(IBW_HUM_01, IBW_HUM_04)), na.rm = TRUE),
         IBW_TEMP_all = rowSums(across(c(IBW_TEMP_01, IBW_TEMP_02, IBW_TEMP_04)), na.rm = TRUE))

# Taped
data_T <- data_T |>
  mutate(IBT_TEMP_all = rowSums(across(c(IBT_TEMP_01, IBT_TEMP_02)), na.rm=TRUE))

# Noise
data_N <- data_N |>
  mutate(NS_all = rowSums(across(c(NS_01)), na.rm = TRUE))

```

```{r plot final2, fig.align='center', fig.width=12, fig.height=2.5, echo=FALSE, message=FALSE, warning=FALSE }
for (uids in unique(data$uid)) {

  plot_H <- data_H |>
    filter(uid == uids) |>
    ggplot(aes(x = datetime)) +
    geom_line(aes(y = IBH_HUM), color = "skyblue2", size = 0.7, alpha = .5) +
    geom_line(aes(y = IBH_TEMP), color = "brown2", size = 0.7, alpha = .5) +
    geom_line(aes(y = ifelse(IBH_HUM_all > 0 ,NA_real_, IBH_HUM)), color = "blue") +
    geom_line(aes(y = ifelse(IBH_TEMP_all > 0 ,NA_real_, IBH_TEMP)), color = "brown") +
    theme_classic() +
    lims(y = c(15, 80)) +
    labs(x = "time", y = "value", title = paste0(uids, " - HOUSE"))

  plot_W <- data_W |>
  filter(uid == uids) |>
  ggplot(aes(x = datetime)) +
    geom_line(aes(y = IBW_HUM), color = "skyblue2", size = 0.7, alpha = .5) +
    geom_line(aes(y = IBW_TEMP), color = "brown2", size = 0.7, alpha = .5) +
    geom_line(aes(y = ifelse(IBW_HUM_all >= 1,NA_real_, IBW_HUM)), color = "blue") +
    geom_line(aes(y = ifelse(IBW_TEMP_all >= 1,NA_real_, IBW_TEMP)), color = "brown") +
    theme_classic() +
    lims(y = c(15, 80)) +
    labs(x = "time", y = "value", title = paste0(uids, " - WORN"))

  plot_T <- data_T |>
    filter(uid == uids) |>
  ggplot(aes(x = datetime)) +
    geom_line(aes(y = IBT_TEMP), color = "brown2", size = 0.7, alpha = .5) +
    geom_line(aes(y = ifelse(IBT_TEMP_all > 0,NA_real_, IBT_TEMP)), color = "brown") +
    theme_classic() +
    labs(x = "time", y = "value", title = paste0(uids, " - TAPED"))

    plot_N <- data_N |>
    filter(uid == uids) |>
    ggplot(aes(x = datetime)) +
    geom_line(aes(y = NS), color = "grey2", size = 0.7, alpha = .5) +
    geom_line(aes(y = ifelse(NS_all >  0, NA_real_, NS)), color = "grey") +
    theme_classic() +
    labs(x = "time", y = "value", title = paste0(uids, " - Noise"))

    grid.arrange(plot_H, plot_W, plot_T, plot_N, ncol=4)
}
```

<br>

### Report on the data
Note: Noise was now averaged with respect to the inherint log-scale.

```{r hourly averages, echo = FALSE, warning=FALSE, message=FALSE}
# House
data_H_hourly <- data_H |>
  mutate(datetime_hour = floor_date(ymd_hms(datetime), "hour"),
         IBH_HUM_clean = if_else(IBH_HUM_all >= 1, NA, IBH_HUM),
         IBH_TEMP_clean = if_else(IBH_TEMP_all >= 1, NA, IBH_TEMP)) |>
  group_by(datetime_hour, uid) |>
  summarise(IBH_HUM_hourly = mean(IBH_HUM_clean, na.rm = TRUE),
            IBH_TEMP_hourly = mean(IBH_TEMP_clean, na.rm = TRUE),
            .groups = "drop") |>
  mutate(uid_hour = paste0(uid, " ", datetime_hour))



# Worn
data_W_hourly <- data_W |>
  mutate(datetime_hour = floor_date(ymd_hms(datetime), "hour"),
         IBW_HUM_clean = if_else(IBW_HUM_all >= 1, NA, IBW_HUM),
         IBW_TEMP_clean = if_else(IBW_TEMP_all >= 1, NA, IBW_TEMP)) |>
  group_by(datetime_hour, uid) |>
  summarise(IBW_HUM_hourly = mean(IBW_HUM_clean, na.rm = TRUE),
            IBW_TEMP_hourly = mean(IBW_TEMP_clean, na.rm = TRUE),
            .groups = "drop") |>
  mutate(uid_hour = paste0(uid, " ", datetime_hour))

# Taped
data_T_hourly <- data_T |>
  mutate(datetime_hour = floor_date(ymd_hms(datetime), "hour"),
         IBT_TEMP_clean = if_else(IBT_TEMP_all >= 1, NA, IBT_TEMP)) |>
  group_by(datetime_hour, uid) |>
  summarise(IBT_TEMP_hourly = mean(IBT_TEMP_clean, na.rm = TRUE),
            .groups = "drop") |>
  mutate(uid_hour = paste0(uid, " ", datetime_hour))

# Noise
data_N_hourly <- data_N |>
  mutate(datetime_hour = floor_date(ymd_hms(datetime), "hour"),
         NS_clean = if_else(NS_all >= 1, NA, NS)) |>
  group_by(datetime_hour, uid) |>
  summarise(NS_hourly = mean(NS, na.rm = TRUE),
            .groups = "drop") |>
  mutate(uid_hour = paste0(uid, " ", datetime_hour))

# combining
data_combined <- data_H_hourly %>%
  full_join(data_W_hourly %>% select(uid_hour, IBW_HUM_hourly, IBW_TEMP_hourly), by = "uid_hour") %>%
  full_join(data_T_hourly %>% select(uid_hour, IBT_TEMP_hourly), by = "uid_hour") %>%
  full_join(data_N_hourly %>% select(uid_hour, NS_hourly), by = "uid_hour")


# diurnal cycle
diurnal_cycle <- data_combined |>
  mutate(hour = as.numeric(format(datetime_hour, "%H"))) |>
  group_by(hour) |>
  summarise(IBH_HUM = mean(IBH_HUM_hourly, na.rm = TRUE),
            IBH_TEMP = mean(IBH_TEMP_hourly, na.rm = TRUE),
            IBW_HUM = mean(IBW_HUM_hourly, na.rm = TRUE),
            IBW_TEMP = mean(IBW_TEMP_hourly, na.rm = TRUE),
            IBT_TEMP = mean(IBT_TEMP_hourly, na.rm = TRUE),
            NS =  10 * log10(mean(10^(NS_hourly/ 10), na.rm = TRUE))) |>
  na.omit()

# Calculate summary statistics (mean, median, sd) for all numeric columns
summary_manual <- data.frame(
  Variable = c("IBH_HUM", "IBH_TEMP", "IBW_HUM", "IBW_TEMP", "IBT_TEMP", "NS"),  
  mean = c(mean(data_combined$IBH_HUM_hourly, na.rm = TRUE), mean(data_combined$IBH_TEMP_hourly, na.rm = TRUE),
           mean(data_combined$IBW_HUM_hourly, na.rm = TRUE), mean(data_combined$IBW_TEMP_hourly, na.rm = TRUE),
           mean(data_combined$IBT_TEMP_hourly, na.rm = TRUE), 10 * log10(mean(10^(data_combined$NS_hourly/ 10), na.rm = TRUE))),
  median = c(median(data_combined$IBH_HUM_hourly, na.rm = TRUE), median(data_combined$IBH_TEMP_hourly, na.rm = TRUE),
             median(data_combined$IBW_HUM_hourly, na.rm = TRUE), median(data_combined$IBW_TEMP_hourly, na.rm = TRUE),
             median(data_combined$IBT_TEMP_hourly, na.rm = TRUE),  10 * log10(median(10^(data_combined$NS_hourly/ 10), na.rm = TRUE))),
  sd = c(sd(data_combined$IBH_HUM_hourly, na.rm = TRUE), sd(data_combined$IBH_TEMP_hourly, na.rm = TRUE),
         sd(data_combined$IBW_HUM_hourly, na.rm = TRUE), sd(data_combined$IBW_TEMP_hourly, na.rm = TRUE),
         sd(data_combined$IBT_TEMP_hourly, na.rm = TRUE),  10 * log10(sd(10^(data_combined$NS_hourly/ 10), na.rm = TRUE))))

```



### Distributions

```{r data distribtuione, echo = FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.height=4, fig.width=8}
plot1 <- ggplot(data_combined, aes(x = IBH_HUM_hourly)) +
  geom_histogram(fill = "skyblue1", color = "black",
                 aes(y = ..density..),
                 bins = 30, 
                 alpha = 0.6) + 
  labs(x = "humidity [%]", y = "density", title = "House - Humidity") +
  lims(x = c(0,100)) +
  theme_classic()


plot2 <- ggplot(data_combined, aes(x = IBW_HUM_hourly)) +
  geom_histogram(fill = "skyblue1", color = "black",
                 aes(y = ..density..),
                 bins = 30, 
                 alpha = 0.6) + 
  labs(x = "humidity [%]", y = "density", title = "Worn - Humidity") +
  lims(x = c(0,100)) +
  theme_classic()


plot3 <- ggplot(data_combined, aes(x = IBH_TEMP_hourly)) +
  geom_histogram(fill = "brown3", color = "black",
                 aes(y = ..density..),
                 bins = 30, 
                 alpha = 0.6) + 
  labs(x = "temperature [°C]", y = "density", title = "House - Temperature") +
  lims(x = c(15,45)) +
  theme_classic()

plot4 <- ggplot(data_combined, aes(x = IBW_TEMP_hourly)) +
  geom_histogram(fill = "brown3", color = "black",
                 aes(y = ..density..),
                 bins = 30, 
                 alpha = 0.6) + 
  labs(x = "temperature [°C]", y = "density", title = "Worn - Temperature") +
  lims(x = c(15,45)) +
  theme_classic()

plot5 <- ggplot(data_combined, aes(x = NS_hourly)) +
  geom_histogram(fill = "grey", color = "black",
                 aes(y = ..density..),
                 bins = 30, 
                 alpha = 0.6) + 
  labs(x = "noise [dB]", y = "density", title = "House - Noise") +
  lims(x = c(15,90)) +
  theme_classic()

plot6 <- ggplot(data_combined, aes(x = IBT_TEMP_hourly)) +
  geom_histogram(fill = "brown3", color = "black",
                 aes(y = ..density..),
                 bins = 30, 
                 alpha = 0.6) + 
  labs(x = "temperature [°C]", y = "density", title = "Taped - Temperature") +
  lims(x = c(15,45)) +
  theme_classic()

grid.arrange(plot1, plot2, plot5, plot3, plot4,  plot6, nrow = 2)

```

### Diurnal Cycle

```{r data diurnal, echo = FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.height=5.5, fig.width=7}
plot1 <- ggplot(diurnal_cycle, aes(x = hour)) + 
  geom_line(aes(y = IBH_HUM, color = "House"), size = 1.2) + 
  geom_line(aes(y = IBW_HUM, color = "Worn"), size = 1.2) + 
  labs(
    x = "hour of the day", 
    y = "humidity [%]", 
    title = "Humidity",
    color = "Legend"  # Title for the legend
  ) +
  lims(y = c(20,50)) +
  scale_color_manual(values = c("House" = "skyblue", "Worn" = "skyblue4"),
                     breaks = c("Worn", "House")) + 
  theme_classic()


plot2 <- ggplot(diurnal_cycle, aes(x = hour)) + 
  geom_line(aes(y = IBH_TEMP, color = "House"), size = 1.2) + 
  geom_line(aes(y = IBW_TEMP, color = "Worn"), size = 1.2) + 
  geom_line(aes(y = IBT_TEMP, color = "Taped"), size = 1.2) + 
  labs(
    x = "hour of the day", 
    y = "temperature [°C]", 
    title = "Temperature",
    color = "Legend"  # Title for the legend
  ) +
  lims(y = c(22,40)) +
  scale_color_manual(
    values = c("House" = "red", "Worn" = "brown3", "Taped" = "brown4"), 
    breaks = c("Taped", "Worn", "House")  # Set the desired legend order
  ) +
  theme_classic()


plot3 <- ggplot(diurnal_cycle, aes(x = hour)) + 
  geom_line(aes(y = NS, color = "Noise"), size = 1.2) + 
  labs(
    x = "hour of the day", 
    y = "noise [dB]", 
    title = "Noise",
    color = "Legend"
  ) +
  lims(y = c(40,70)) +
    scale_color_manual(
    values = c("Noise" = "darkgrey")) +
  theme_classic()

grid.arrange(plot1, plot2, plot3, ncol = 1)
```

### Descriptive Statistics

```{r table stats, echo = FALSE}

kable(summary_manual, 
      format = "markdown",  
      col.names = c("Variable", "Mean", "Median", "Standard Deviation"))
```

### Scatterplots

```{r intra device, echo = FALSE, warning=FALSE, messages = FALSE, fig.align='center', fig.width=11, fig.height=4}

plot1 <- ggplot(data_combined, aes(x = IBH_TEMP_hourly, y = IBH_HUM_hourly)) +
  geom_point(size = 0.7, color = "purple1") +
  labs(x = "temperature [°C]", y = "humidity [%]", title = "House") +
  theme_classic()

plot2 <- ggplot(data_combined, aes(x = IBW_TEMP_hourly, y = IBW_HUM_hourly)) +
  geom_point(size = 0.7, color = "purple1") +
  labs(x = "temperature [°C]", y = "humidity [%]", title = "Worn") +
  theme_classic()

plot3 <- ggplot(data_combined, aes(x = IBH_HUM_hourly, y = IBW_HUM_hourly)) +
  geom_point(size = 0.7, color = "skyblue2") +
  labs(x = "humidity [%]", y = "humidity [%]", title = "House x Worn") +
  theme_classic()

grid.arrange(plot1, plot2, plot3, nrow = 1)
```

<br>

```{r scatterl interdevice, echo=FALSE, fig.align='center', fig.height=6, fig.width=7, warning=FALSE, message=FALSE}
plot4 <- ggplot(data_combined, aes(x = IBH_TEMP_hourly, y = IBT_TEMP_hourly)) +
  geom_point(size = 0.7, color = "brown2") +
  labs(x = "temperature [°C]", y = "temperature [°C]", title = "House x Taped") +
  theme_classic()


plot5 <- ggplot(data_combined, aes(x = IBW_TEMP_hourly, y = IBT_TEMP_hourly)) +
  geom_point(size = 0.7, color = "brown2") +
  labs(x = "temperature [°C]", y = "temperature [°C]", title = "Worn x Taped") +
  theme_classic()



plot6 <- ggplot(data_combined, aes(x = IBW_TEMP_hourly, y = IBH_TEMP_hourly)) +
  geom_point(size = 0.7, color = "brown2", alpha = 0.7) +
  labs(x = "temperature [°C]", y = "temperature [°C]", title = "Worn x House") +
  theme_classic()

plot7 <- ggplot(data_combined, aes(x = IBH_TEMP_hourly, y = NS_hourly)) +
  geom_point(size = 0.7, color = "#884848") +
  labs(x = "temperature [°C]", y = "noise [dB]", title = "House") +
  theme_classic()


grid.arrange(plot4, plot5,  plot6, plot7, nrow = 2)
```

<br>